---
title: "Introduction to Dimensional Reduction in R - UseR!22 (Section IV)"
author: "Isabella Bicalho Frazeto"
date: '2022-06-19'
output: html_document
---

# Section 4


## Import Reticulate 

```{r}
library(reticulate)
```

We have to check that we have indeed create an environment link:

```{r}
Sys.getenv("RETICULATE_PYTHON")
```
If the string is empty, do not panic!
There are some links to solve this:  1) https://solutions.rstudio.com/r/reticulate/
2) https://docs.python.org/3/library/venv.html
  
For me it looks like this:
  1)  python3 -m venv .venv
  2) echo "RETICULATE_PYTHON=/usr/bin/python3.6" >> .Renviron
  3)  echo ".Renviron" >> .gitignore


# Using python 

```{python}
a = "Hello World" + " from Python"
print(a)

```

# Calling a variable from python in R
```{R}
py$a
```

# How to call a variable from R in python
```{R}
b <- "Hello World from R"
```

```{python}
print(r.b)
```

## We can start by making a dataset that has non-linear relationship to it

```{python}
import matplotlib.pyplot as plt
from sklearn import manifold, datasets

sr_points, sr_color = datasets.make_swiss_roll(n_samples=1500, noise = 1, random_state=0)
```

## Showing our roll from python 

```{python}

fig = plt.figure(figsize=(8, 6))
ax = fig.add_subplot(111, projection="3d")
fig.add_axes(ax)
ax.scatter(
    sr_points[:, 0], sr_points[:, 1], sr_points[:, 2], c=sr_color, s=50, alpha=0.8
)
ax.set_title("Swiss Roll in Ambient Space")
ax.view_init(azim=-66, elev=12)
_ = ax.text2D(0.8, 0.05, s="n_samples=1500", transform=ax.transAxes)


```

## Using T-SNE

```{python}
sr_tsne = manifold.TSNE(
    n_components=2, perplexity=40, init="random", random_state=0
)
```

TIP: You can access the documentation in python by calling "py_help(py$manifold)".

# Fitting the T-SNE

```{python}
fitted_tsne = sr_tsne.fit_transform(sr_points)
```


# Showing our T-SNE in R from python

```{r}
library(ggplot2)
library(tidyverse)
library(cowplot)

py$fitted_tsne %>%
  as_tibble()  %>%
  ggplot(aes(x=V1, y=V2,  color = py$sr_color)) +
  geom_point() +
  labs(color = "Color") +
  theme_minimal_grid() +
  panel_border(color = "black") +
  scale_color_continuous(type = "viridis")
```

# UMAP from UMAP

```{R}
library(umap)
umap_learn = umap(py$sr_points)

```

Seeing our results:

```{r}
umap_tibble <- umap_learn$layout %>% 
                as_tibble() %>%
                mutate(color= py$sr_color)

umap_tibble
```


```{r}
umap_tibble %>%
  ggplot(aes(x=V1, y=V2, color = color )) +
  geom_point() +
  theme_minimal_grid() +
  panel_border(color = "black") +
  scale_color_continuous(type = "viridis")
```


