---
title: "Introduction to Dimensional Reduction in R - UseR!22 (Section III)"
author: "Isabella Bicalho Frazeto"
date: '2022-06-19'
output: html_document
---

# Section 3 

## Imports
```{r}
library(tidyverse)
library(ggplot2)
library(vegan)
library(MASS)
library(ggforce)
library(cowplot)
```

## Getting to know our dataset

```{r}

data("swiss")
head(swiss)
skimr::skim(swiss)

```

```{r}
swiss %>%
  ggplot(aes(x=Education, y = Examination)) +
  geom_point() +
  theme_minimal_grid() +
  panel_border(color = "black")

```


```{r}
swiss %>%
  ggplot(aes(x=Fertility, y = Agriculture)) +
  geom_point() +
  theme_minimal_grid() +
  panel_border(color = "black")
```



# Base R
We can calculate the Principal Coordinate Analysis (i.e Metric Multidimensional Scaling or Classical multidimensional scaling) using the cmdscale function in R.

```{r}
distance_matrix_euclidean <- swiss %>% dist(method = "euclidean")

euclidean_MDS <- distance_matrix_euclidean %>% 
                  cmdscale()

tibble_euclidean_MDS <- tibble(
                          MDS_1 = euclidean_MDS[,1], 
                          MDS_2 = euclidean_MDS[,2],
                          city =  rownames(euclidean_MDS)
)
                  

tibble_euclidean_MDS %>%
  ggplot(aes(x=MDS_1, y = MDS_2, label = city)) +
  geom_text(check_overlap = TRUE) +
  theme_minimal_grid() +
  panel_border(color = "black")

```


## Let's do it using another distance 

```{r}

```


## The vegan package also has a similar fuction (metaMDS)
```{r}
library(vegan)
vegan_mds <- metaMDS(comm = swiss, 
                     distance = "jaccard", 
                     trace = FALSE,
                     autotransform = FALSE)

```

Similarly, we can plot it.

```{r}
plot(vegan_mds$points)

```

It is not super informative. We can make it better. Let's start by creating a dataframe

```{r}

df_MDS <- data.frame(vegan_mds$points)

```


We need to make sure our names are matching 

```{r}
#sanity check
rownames(df_MDS) == rownames(swiss)
```

```{r}
df_MDS$Education <- swiss$Education
df_MDS$Fertility <- swiss$Fertility
df_MDS$inf_mort <- swiss$Infant.Mortality

df_MDS %>% 
  ggplot(aes(x=MDS1, y=MDS2, color =  Fertility, shape = Education > 20)) +
  geom_point(size = 4) +
  theme_minimal_grid() +
  panel_border("black") +
  scale_color_continuous(type = "viridis")
```

## Education

```{r}

df_MDS %>% 
  ggplot(aes(x=MDS1, y=MDS2, color = Education)) +
  geom_point() +
  theme_minimal_grid() +
  panel_border(color="black") +
  scale_color_continuous(type = "viridis")

```

## What about infant mortality?

```{r}
df_MDS %>% 
  ggplot(aes(x=MDS1, y=MDS2, label = rownames(df_MDS), color =  inf_mort)) +
  geom_text(check_overlap = TRUE) +
  theme_minimal_grid() +
  panel_border(color = "black")

```

## Finally, we can check the stress 

```{r}

vegan_mds$stress
```

### Non-parametric multidimensional scaling 

```{r}
library(MASS)

non_metric_mds <- swiss %>%
                    dist("canberra") %>%          
                    isoMDS(k=9) 

```

Let's explore our results. We will extract the point and pass it as data.frame

```{r}

non_metric_mds_df <- non_metric_mds$points
colnames(non_metric_mds_df) <- c(paste0("DIM", 1:9))
non_metric_mds_df

```

We also plot the dimensions like we did for PCA and ICA

```{r}
non_metric_mds_df %>%
  as_tibble() %>%
  mutate(
    city = rownames(non_metric_mds_df)
  ) %>%
  ggplot(aes(x = .panel_x, y = .panel_y)) +
  geom_point(alpha = 0.4, size = 0.5) +
  geom_autodensity(alpha = .3) +
  facet_matrix(vars(-city), layer.diag = 2) + 
  scale_color_brewer(palette = "Dark2") + 
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal_grid() +
  panel_border(color = "black") +
  theme(legend.position = "bottom")
```

Is there something else you would like to here? Another k? Maybe another distance? Who wants to try another dataset?